---
title: "Digital epigraphy in 2022: state of the art"
description: 
author: 
 - Petra Hermankova^[Johannes Gutenberg University in Mainz, petra.hermankova@uni.mainz.de, https://orcid.org/0000-0002-6349-0540]
 - Marietta Horster^[Johannes Gutenberg University in Mainz, horster@uni.mainz.de, https://orcid.org/0000-0003-1434-224X ]
 - Jonathan Prag^[Oxford University, jonathan.prag@merton.ox.ac.uk, https://orcid.org/0000-0003-3819-8537 ]
date: "28/2/2022"
output: 
  html_document:
    css: ../"style.css"
    theme: paper
    toc: yes
    toc_float: true
    number_sections: true
    smooth_scroll: true
    toc_depth: 4
    df_print: paged
    highlight: monochrome
    fontsize: 12pt
    
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
```

# Introduction
The aim of the report is to summarise the state of digital epigraphy in 2022 and to do so in a transparent and FAIR manner. This report has been compiled on the basis of an online survey *FAIR Epigraphy: Scoping survey for partners and collaborators* and *Digital epigraphy in 2022: scoping survey* for all digital epigraphy projects, circulating in February 2022. 

The survey was organised by The FAIR Epigraphy Project (https://www.csad.ox.ac.uk/fair-epigraphy), a collaboration between Johannes Gutenberg University in Mainz (Prof. Marietta Horster) and Oxford University (Prof. Jonathan Prag). The project is funded by the Arts and Humanities Research Council (AHRC) and Deutsche Forschungemeinschaft (DFG) and will run for 36 months from 2022-2025.

The FAIR Epigraphy project was launched in February 2022, aiming to create an interactive platform for all epigraphic projects, aligning their digital needs with the principles of FAIR science. *"The principles emphasise machine-actionability (i.e., the capacity of computational systems to find, access, interoperate, and reuse data with none or minimal human intervention) because humans increasingly rely on computational support to deal with data as a result of the increase in volume, complexity, and creation speed of data."* (Quote from the FAIR Principles website, https://www.go-fair.org/fair-principles/) 
The FAIR Epigraphy project does not wish to replicate any current efforts, but align existing initiatives and bring them together to create a hub of high-quality tools and FAIR compliant standards and resources for the modern epigraphic discipline.

In order to map the existing field of digital epigraphy, current practices and standards, as well as find out the (digital) needs of the discipline, we have circulated the two scoping surveys in February 2022. The results of the survey, presented in the current report, will be used to plan the activities and efficiently allocate the resources of the FAIR Epigraphy Project in the next three years (2022-2024). The survey answers are anonymised so the individual projects cannot be identified on the basis of their answers, and stored as .TSV (tab separated values) within the same GitHub repository (https://github.com/FAIR-epigraphy/scoping_survey_report) as a supplement to the text of the report.

# FAIR Epigraphy partner projects

```{r data_partners, include=FALSE}
# loading data from TSV file
partners<- read_tsv("../data/01_scoping_survey_anonymised.tsv")
```

This section summarises the results of the online survey *FAIR Epigraphy: Scoping survey for partners and collaborators* aimed at established digital projects that became official partners and collaborators with the FAIR Epigraphy Project.
Originally there were 14 partner projects. We have recieved `r nrow(partners)` responses to the survey, with the response rate `r round(nrow(partners)/0.14, 2)`%.

The partner projects represent relatively established projects with the average duration of a project shortest being `r median(partners$Length_numeric_years)` year(s). The shortest participating project reported their duration as `r min(partners$Length_numeric_years)` year(s) and the longest `r max(partners$Length_numeric_years)` year(s).

## Language coverage

**Question:** *What is the predominant language of epigraphic data in your project (for mixed collections or collections where other languages are predominant provide details in Other)*
```{r}
partners_language <- as.data.frame(unlist(str_split(partners$Language_clean, "; ")))
colnames(partners_language) <- c("language")

partners_language_ratio <- partners_language %>% 
  count(language, sort=T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
partners_language_ratio
```
**Commentary:** The language coverage of the participating projects consisted predominantly from Latin and Greek projects (representing `r sum(partners_language_ratio$ratio[1:2])`% of the answers). The Other category encompassed a substantial part of the surveyed projects, documenting the need to expand beyond the traditional Latin and Greek focus of epigraphic discipline. The languages listed as `Other` consisted of `r partners_language_ratio$language[4:10]`.

## IT infrastructure

**Question**: *Does the project have a website?*
```{r}
partners %>% 
  count(Website)
```
**Commentary**: All of the participating projects maintain an online presence (as of February 2022).

---

**Question**: *Does you project have an IT specialist(s)?*
```{r}
IT <- partners %>% 
  count(`Does you project have an IT specialist(s)?`, sort=T) %>% 
  rename(IT_spec = `Does you project have an IT specialist(s)?`) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
IT
```
**Commentary**: Vast majority of digital projects have an IT specialist, yet only `r sum(IT$n[3:4])` projects have an equivalent of 1.0 FTE or more to their disposal. The `r round(IT$ratio[1],2)` % of projects have access to part-time IT support for their projects, which in some instances may be only a few hours per week per project.

---

**Question**: *Does your project store epigraphic data in the following formats...?* 
```{r}
format <- partners %>% 
  count(`Our projects stores epigraphic data in the following formats`, sort=T) %>% 
  rename(format = `Our projects stores epigraphic data in the following formats`) %>% 
  mutate(no_format = str_count(format, ",") + 1)
format
```
**Commentary**: Majority of projects use Epidoc XML as their main output data format (representing `r round(sum(format$n[1:4])/(nrow(partners)/100),2)`% of participating projects). `r round((sum(format$no_format == 1))/(nrow(partners)/100),2)`% of projects use only one type of data format, while `r round((sum(format$no_format > 1))/(nrow(partners)/100),2)`% two or more data format types (such as `r unique(unlist(str_split(format$format, ", ")))`).


## Data sharing

**Question**: Do you share your data outside of your project?
```{r}
share<- partners %>% 
  count(`Do you share your data outside of your project`, sort=T) %>% 
  rename(share = `Do you share your data outside of your project`) %>% 
  mutate(ratio = round((n/(sum(n)/100)),2))
share
```
**Commentary**: Majority of partner projects share their data: `r share$ratio[1]`% shares the data under a Creative Common license, which is the preferred more according to the FAIR data principles. All partner projects reported their willingness to share the data, even if they are not currently doing it, or provide the data only on demand.

---

**Question**: *How do share your data with users outside your project?*
```{r}
outshare<- partners %>% 
  count(`How do share your data with users outside your project`, sort=T) %>% 
  rename(share_all = `How do share your data with users outside your project`) %>% 
  mutate(share_method = str_count(share_all, "; ") + 1) 

outshare
```
**Commentary**: All partner projects provide at least one way of sharing the data (whether it maybe currently accesible to public or not, or it is intended to be accessible in the future). The average (median) number of sharing methods per project is `r median(outshare$share_method)`.

---

**Question**: *What is the number of sharing methods accross projects?*
```{r, fig.width=10}
sharingserv <- as.data.frame(unlist(str_split(partners$`How do share your data with users outside your project`, "; ")))
colnames(sharingserv) <- c("service")


sharing_service <- sharingserv %>% 
  count(service, sort= T) 

sharingplot<- sharing_service %>% 
  ggplot() + 
  geom_col(aes(y= reorder(service, -n), x= n, fill = n), width = 0.9) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 75))  +
  ggtitle("Number of sharing methods per project") +
  xlab("Sharing method") + ylab("Number of projects") +
  #coord_flip() +
  theme_bw() +
  #coord_cartesian(xlim = c(0, 5)) +
  theme(axis.text = element_text(size = 9)) +
  theme(plot.title = element_text(size = 13))

sharingplot
```
**Commentary**: Epidoc XML is by far the most popular format for data sharing, however other Open Science services are starting to make it way to digital epigraphy, such as sharing via public repository on GitHub or Zenodo, as well as providing raw data in the CSV (comma separated value) format, or JSON files. Only a relative minority of participating partner projects shares the data on demand or have a non-public API access point to their data.

## Institutional policies

**Question:** *Does your institution of funding body require your project to comply with any data policies (e.g., FAIR principles, data storage, data sharing, Open Science)?*

```{r}
policies <- partners %>% 
  mutate(policies = `Does your institution of funding body require your project to comply with any data policies (e.g., FAIR principles, data storage, data sharing, Open Science)?`) %>% 
  count(policies, sort=T) %>% 
  mutate(ratio = round((n/(sum(n)/100)),2))
policies
```
**Commentary**: The majority of projects (`r policies$ratio[1]`%) are required to comply with data-related policy introduced either by their institution or funding body. Only a minority of partner projects (`r policies$ratio[2]`%) are not required to follow nay data policy, nor follow it on a voluntary basis.

---

**Question**: *If you have answered YES in the previous question, please specify what are the policies, or provide link.*

```{r}
unique(na.omit(partners$`If you have answered YES in the previous question, please specify what are the policies, or provide link.`))
```
**Commentary**: Several of the partner projects follow the ERC data and open access policies. More information on ERC Open Research Data and Data Management Plans can be found at https://erc.europa.eu/sites/default/files/document/file/ERC_info_document-Open_Research_Data_and_Data_Management_Plans.pdf or at ERC Open Science policies page https://erc.europa.eu/managing-your-project/open-science.

## Open Science practice

**Question**: *Standardized terminologies: The project uses the following systems:*
```{r}
stand_term <- as.data.frame(unlist(str_split(partners$`Standardized terminologies: The project uses the following systems (check all that apply)`, ", ")))
colnames(stand_term) <- c("standard_terminologies")

stand_term_ratio<- stand_term %>% 
  count(standard_terminologies, sort= T) %>% 
  mutate(ratio = round(n/(nrow(partners)/100),2))  

stand_term_ratio
```
**Commentary**: The lists of vocabularies for the epigraphic discipline created by the EAGLE project (https://www.eagle-network.eu/resources/vocabularies/) are used by most of the projects: either in the original form (representing `r stand_term_ratio$ratio[3]`% of participating projects) or in the form modified for the needs of the project (representing `r stand_term_ratio$ratio[1]`% of participating projects). The need of modifications suggests that the EAGLE vocabularies do not in fact form a community wide standard and need to be improved before becoming one. 

---

**Question**: *Standardized terminologies: data on combination of vocabularies systems* 
```{r}
stand_method<- partners %>% 
  #count(`Standardized terminologies: The project uses the following systems (check all that apply)`, sort=T) %>% 
  rename(stand_all = `Standardized terminologies: The project uses the following systems (check all that apply)`) %>% 
  mutate(standard_method_no = str_count(stand_all, ", ") + 1)  %>% 
  select(standard_method_no) %>% 
  count(standard_method_no) %>% 
  mutate(ratio = round(n/(nrow(partners)/100),2))
stand_method
```
**Comentary**: Majority of projects (`r stand_method$ratio[1]`%) uses only one method to record their standards, while `r sum(stand_method$ratio[2:3])`% of projects use a combination of two or three methods. The internal authority lists are used in combination with the EAGLE vocabularies in its original and modified form. Sharing of internal authority lists would be thus highly beneficial for improving the existing EAGLE vocabularies.

---

**Question**: *Linked Open Datasets: The project uses the following systems:*
```{r}
lod <- as.data.frame(unlist(str_split(partners$`Linked Open Data: The project uses the following systems (check all that apply)`, ", ")))
colnames(lod) <- c("linked_data")


lod_ratio<- lod %>% 
  count(linked_data, sort= T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2)) %>% 
  mutate(ratio_all_proj = round(n/(nrow(partners)/100),2))
lod_ratio
```
**Commentary**: From the listed Linked Open Datasets (LOD), Pleaides and Trismegistos are by far the most popular, being used in `r lod_ratio$ratio_all_proj[1]`% of all participating projects. EAGLE vocabularies are used in `r lod_ratio$ratio_all_proj[3]`% of all participating projects. The onomastic and prosopographic data, represented by EDH People, PIR, and LGPN are used only sporadically, namely in `r lod_ratio$ratio_all_proj[4]`% or `r lod_ratio$ratio_all_proj[5]`% of all participating projects, suggesting there is a great space for improvement and potentially great benefit in creating onomastic and prosopographic LOD.

# Digital epigraphy projects
```{r data_projects, include=FALSE}
# loading data from TSV file
projects<- read_tsv("../data/02_scoping_survey_anonymised.tsv")
```
This section summarises the results of the online survey *Digital epigraphy in 2022: scoping survey* aimed at digital projects listed under the Digital Epigraphy Projects on the Digital Classicist Wiki page (https://wiki.digitalclassicist.org/Category:Projects). Originally the survey was sent to 83 projects. We have recieved `r nrow(projects)` responses to the survey, with the response rate `r round(nrow(projects)/0.83, 2)`%.

The digital projects represent a wide range of projects from well established projects to short-term mostly PhD projects, with the average duration of a project shortest being `r median(projects$Length_numeric_years)` year(s). The shortest participating project reported their duration as `r min(projects$Length_numeric_years)` year(s) and the longest `r max(projects$Length_numeric_years)` year(s).

**Question**: *Is the project still active?*
```{r}
proj_act<- projects %>% 
  count(`Is the project still active?`) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
proj_act
```
**Commentary**: `r proj_act$ratio[3]`% of participating projects are still active, while `r proj_act$ratio[2]`% of projects are permanently closed and do not consider restarting in the future. `r proj_act$ratio[1]`% of projects are currently not active, but might be reactivated in the future. 

## Language coverage

**Question:** *What is the predominant language of epigraphic data in your project (for mixed collections or collections where other languages are predominant provide details in Other)*
```{r}
projects_language <- as.data.frame(unlist(str_split(projects$Language_clean, "; ")))
colnames(projects_language) <- c("language")

projects_language_ratio <- projects_language %>% 
  count(language, sort=T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
projects_language_ratio
```
**Commentary:** The language coverage of the participating projects consisted predominantly from Latin and Greek projects (representing `r sum(projects_language_ratio$ratio[1:2])`% of the answers). The Other category encompassed a substantial part of the surveyed projects, documenting the need to expand beyond the traditional Latin and Greek focus of epigraphic discipline. The languages listed as `Other` consisted of `r projects_language_ratio$language[4:15]`.


## IT infrastructure

**Question**: *Does the project have a website?*
```{r}
projects %>% 
  count(Website)
```
**Commentary**: The majority of the participating projects maintain an online presence (as of February 2022).

---

**Question**: *Does you project have an IT specialist(s)?*
```{r}
IT2 <- projects %>% 
  count(`Does you project have an IT specialist(s)?`, sort=T) %>% 
  rename(IT_spec = `Does you project have an IT specialist(s)?`) %>% 
  #na.omit(IT_spec) %>% 
  mutate(ratio = round(n/(nrow(projects)/100),2))
IT2
```
**Commentary**: The `r sum(IT2$ratio[c(2,5,6,8)])`% of digital projects have an IT specialist available for at least several hours, yet only `r IT2$ratio[4]`% of projects have an equivalent of 1.0 FTE or more to their disposal. The `r sum(IT2$ratio[c(2,6,8)])` % of projects have access to part-time IT support for their projects, which in some instances may be only a few hours per week per project. Several projects report difficulty with finding financial resources to support further development, or day-to-day support. The `r sum(IT2$ratio[c(2,7)])` % of projects report currently not having any access to IT support. The `r IT2$ratio[1]`% projects did not indicate whether they have access to IT support.

---

**Question**: *Does your project store epigraphic data in the following formats...?* 
```{r}
format2 <- projects %>% 
  count(`Our projects stores epigraphic data in the following formats`, sort=T) %>% 
  rename(format = `Our projects stores epigraphic data in the following formats`) %>% 
  mutate(no_format = str_count(format, ",") + 1)
format2
```
**Commentary**: A large portion of projects did not indicate any format of the data  `r round(sum(format2$n[c(3,5,6,7)])/(nrow(projects)/100),2)`%. Epidoc XML represented one the two main output data formats with `r round(sum(format2$n[c(3,5,6,7)])/(nrow(projects)/100),2)`% of participating projects, while the data stored in SQL and similar database languages represented `r round(sum(format2$n[c(2,4,7,8,10)])/(nrow(projects)/100),2)`% of participating projects. `r round(format2$n[10]/(nrow(projects)/100),2)`% projects indicated use of analogue data, such as printed materials and 3D format. `r round(format2$n[11]/(nrow(projects)/100),2)`% projects indicated using their own version of Epidoc XML, adapted to their specific needs. `r round((sum(na.omit(format2$no_format == 1)))/(nrow(projects)/100),2)`% of projects use only one type of data format, while `r round((sum(na.omit(format2$no_format > 1)))/(nrow(projects)/100),2)`% two or more data format types (such as `r unique(unlist(str_split(format2$format, ", ")))[2:6]`).


## Data sharing

**Question**: Do you share your data outside of your project?
```{r}
share2<- projects %>% 
  count(`Do you share your data outside of your project`, sort=T) %>% 
  rename(share = `Do you share your data outside of your project`) %>% 
  mutate(ratio = round((n/(sum(n)/100)),2))
share2
```
**Commentary**: Large part of projects did not indicate whether they share their data (`r share2$ratio[1]`%). However, the majority of projects are willing to their their data, representing `r sum(share2$ratio[c(2,4,5,6,7,8,9,10,11)])`% of participating projects. `r sum(share2$ratio[c(2,8,9,10)])`% of participating projects share the data under a Creative Common license, which is the preferred more according to the FAIR data principles. `r sum(share2$ratio[c(4,11)])`% of participating projects share the data without any specific license, `r share2$ratio[5]`% of participating projects provide the data only on demand.

---

**Question**: *How do share your data with users outside your project?*
```{r}
outshare2<- projects %>% 
  count(`How do share your data with users outside your project`, sort=T) %>% 
  rename(share_all = `How do share your data with users outside your project`) %>% 
  mutate(share_method = str_count(share_all, "; ") + 1) 

outshare2
```
**Commentary**: All partner projects provide at least one way of sharing the data (whether it maybe currently accesible to public or not, or it is intended to be accessible in the future). The average (median) number of sharing methods per project is `r median(na.omit(outshare2$share_method))`.

---

**Question**: *What is the number of sharing methods accross projects?*
```{r, fig.width=10}
sharingserv2 <- as.data.frame(unlist(str_split(projects$`How do share your data with users outside your project`, "; ")))
colnames(sharingserv2) <- c("service")


sharing_service2 <- sharingserv2 %>% 
  count(service, sort= T) 

sharingplot2<- sharing_service2 %>% 
  ggplot() + 
  geom_col(aes(y= reorder(service, -n), x= n, fill = n), width = 0.9) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))  +
  ggtitle("Number of sharing methods per project") +
  xlab("Sharing method") + ylab("Number of projects") +
  #coord_flip() +
  theme_bw() +
  coord_cartesian(xlim = c(0, 7)) +
  theme(axis.text = element_text(size = 7)) +
  theme(plot.title = element_text(size = 13))

sharingplot2
```
**Commentary**: Majority of digital projects does not provide any indication of sharing their data, which severely limits any attempts of dicispline-wide integration of data. From those who share the data, the Epidoc XML format is the most popular format for data sharing, as well as seach output on the project's website. Open Science practices do not seem to be popular choice in digital epigraphy, such as sharing via public repository on GitHub or Zenodo, as well as providing raw data in the CSV (comma separated value) format, or JSON files. Computer-automated-acess to data is rare and manual human interaction, such as manual selection, manual download of files prevails, potentially hindering any quantitative and reproducible studies, or linking of datasets via automating processes. For example, an API access point is currently available only for 1 project.


#Petra continue here -----


### Institutional policies

*Question:* Does your institution of funding body require your project to comply with any data policies (e.g., FAIR principles, data storage, data sharing, Open Science)?

```{r}
projects %>% 
  mutate(policies = `Does your institution of funding body require your project to comply with any data policies (e.g., FAIR principles, data storage, data sharing, Open Science)?`) %>% 
  count(policies, sort=T)
```

**Question**: If you have answered YES in the previous question, please specify what are the policies, or provide link.

```{r}
unique(na.omit(projects$`If you have answered YES in the previous question, please specify what are the policies, or provide link.`))
```

### FAIR Principles

#### Standardized terminologies

The project uses the following systems:
```{r}
stand_term2 <- as.data.frame(unlist(str_split(projects$`Standardized terminologies: The project uses the following systems (check all that apply)`, ", ")))
colnames(stand_term2) <- c("standard_terminologies")


stand_term2 %>% 
  count(standard_terminologies, sort= T) %>% 
  mutate(ratio = n/(sum(n)/100))
```


#### Linked Open Datasets

The project uses the following systems:
```{r}
lod2 <- as.data.frame(unlist(str_split(projects$`Linked Open Data: The project uses the following systems (check all that apply)`, ", ")))
colnames(lod2) <- c("linked_data")


lod2 %>% 
  count(linked_data, sort= T) %>% 
  mutate(ratio = n/(sum(n)/100))
```

# Future needs of digital epigraphy

This section covers the wishes of partner projects as well as all participating digital epigraphy projects. The responses were anonymised so no individual or project can be identified but otherwise presented as submitted in the survey.

## Partner projects 

**Question**: *Our project would like to be able to use within the next three years:*
```{r}
lod_future <- as.data.frame(unlist(str_split(partners$`Select all options your project like to be able to use within the next three years:`, ", ")))
colnames(lod_future) <- c("lod_f")


lod_f_ratio<- lod_future %>% 
  count(lod_f, sort= T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2)) %>% 
  mutate(ratio_all_proj = round(n/(nrow(partners)/100),2))
lod_f_ratio
```
**Commentary**: The most popular is the option `Bibliographical references to all epigraphic publications with stable URI (e.g. Zenon)` with `r lod_f_ratio$ratio[1]`% of responses representing the wishes of `r lod_f_ratio$ratio_all_proj[1]`% of all projects. The great interest in onomastic and prosopographical LOD for both Greek and Roman world is supported by `r sum(lod_f_ratio$ratio[c(3:4)])`% of positives responses from `r lod_f_ratio$ratio_all_proj[3]`% and `r lod_f_ratio$ratio_all_proj[4]`% of projects respectively. The `improved EAGLE vocabularies` are wished for by `r lod_f_ratio$ratio_all_proj[2]`% of participating projects. The `domain specific repository for epigraphic data` or the `open and accessible RDF Triplestore` do not seem to be the highest priority of participating projects, but still relatively popular as `r lod_f_ratio$ratio_all_proj[5]`% of responses wishes for one of the two. One participating project wishes specifically for the following: ``r lod_f_ratio$lod_f[7]``.

--- 
 
**Question**: *Potential ideas that our project would benefit from:*
```{r}
lod_ideas <- as.data.frame(unlist(str_split(partners$`Select all items your project would benefit from in the next three years:`, ", ")))
colnames(lod_ideas) <- c("lod_i")


lod_i_ratio<- lod_ideas %>% 
  count(lod_i, sort= T) %>% 
  mutate(ratio = n/(sum(n)/100)) %>% 
  mutate(ratio_all_proj = round(n/(nrow(partners)/100),2))

lod_i_ratio
```
**Commentary**: `r lod_i_ratio$ratio_all_proj[1]`% of all projects would benefit from `Set of guidelines for FAIR and Linked Open Data in epigraphy`. There is general interest in practical examples and workshop on how to used LOD and FAIR Principles in Epigraphy, as well as resources for quantitative analysis of data in epigraphy.

---

**Question**: *Additional digital needs*
```{r}
unique(na.omit(partners$`Your personal wish list. Provide a description of potential tasks and issues the FAIR Epigraphy can help your project within the next three years (and beyond)`))
```
**Commentary**: This section covers additional needs of partner projects. 

## Digital epigraphy projects

**Question**: *Our project would like to be able to use within the next three years:*
```{r}
lod_future2 <- as.data.frame(unlist(str_split(projects$`Select all options your project like to be able to use within the next three years:`, ", ")))
colnames(lod_future2) <- c("lod_f")


lod_future2 %>% 
  count(lod_f, sort= T) %>% 
  mutate(ratio = n/(sum(n)/100))
```

**Question**: *Potential ideas that our project would benefit from:*
```{r}
lod_ideas2 <- as.data.frame(unlist(str_split(projects$`Select all items your project would benefit from in the next three years:`, ", ")))
colnames(lod_ideas2) <- c("lod_i")


lod_ideas2 %>% 
  count(lod_i, sort= T) %>% 
  mutate(ratio = n/(sum(n)/100))
```

**Question**: *Additional digital needs*
```{r}
unique(na.omit(projects$`Your personal wish list. Provide a description of potential tasks and issues the FAIR Epigraphy can help your project within the next three years (and beyond)`))
```

